First steps towards a formalization of Forcing

### 2018-08-15 LSFA/rs.tex
más detalle

Two different versions of $\DC$ (called $\DC0$ and $\DC(\kappa)$) have
already been formalized  by Krzysztof 
Grabczewski~\cite{DBLP:journals/jar/PaulsonG96}, as part of a
study of equivalents to $\AC$ (following Rubin and Rubin
\cite{rubin1985equivalents}). Nevertheless, those are not convenient
for our purposes for the reasons we now explain. The axiom $\DC0$
corresponds essentially to our Pointed $\DC$ but without the
constraint $f(0)=a$; it is a nice exercise to show that $\DC0$ implies
Pointed $\DC$, but a formalization would have a moderate length. On
the other hand, the shortest path from $\AC$ as stated in the
corresponding Isabelle theory to $\DC0$ using the already formalized
material involves a complicated detour (130+ proof commands spanning
various files of the \isatt{ZF-AC} theory), compared to the mere 11 commands from $\AC$ to
\isatt{AC{\isacharunderscore}func{\isacharunderscore}Pow}. This last
one is the
choice principle that we use in our formalization of Pointed $\DC$, and states 
the existence of
choice functions on $\P(A)\sm \{\mty\}$):
\[
\exists (s: \P(A)\sm \{\mty\}\to A). \, \forall X\sbq A.\ X\neq \mty \limp
s(X) \in X.
\]
Another advantage of taking
\isatt{AC{\isacharunderscore}func{\isacharunderscore}Pow} as a
starting point is that it does not involve proper classes: The version
of  $\AC$ in Isabelle/ZF corresponds to an axiom scheme of first order
logic and as such is not a standard formulation. 
%% 26 commands to get Pointed $\DC$.

@@

The proof is mostly routine; after a few lemmas (having 26 proof
commands in total) we obtain the
following theorem:

>>>>

Two different versions of $\DC$ (called $\DC0$ and $\DC(\kappa)$) have
already been formalized  by Krzysztof 
Grabczewski~\cite{DBLP:journals/jar/PaulsonG96}, as part of a
study of equivalents to $\AC$ (following Rubin and Rubin
\cite{rubin1985equivalents}). Nevertheless, those are not convenient
for our purposes for reasons we now explain. The axiom $\DC0$
corresponds essentially to our Pointed $\DC$ but without the
constraint $f(0)=a$; it is a nice exercise to show that $\DC0$ implies
Pointed $\DC$, but a formalization would have a moderate
length. $\DC(\kappa)$ is rather different in nature and it is tailored
to  obtain another proposition
equivalent to the axiom of choice (actually,
$\AC\longleftrightarrow (\forall \kappa
.\; \mathrm{Card}(\kappa)\longrightarrow \DC(\kappa))$. 
Finally, the shortest path from $\AC$ %% as stated in the 
%% corresponding Isabelle theory
to $\DC0$ using already formalized
material involves a complicated detour (130+ proof commands spanning
various files of the \isatt{ZF-AC} theory and going through the Well
Ordering Theorem and  $\DC(\omega)$), compared to the mere 11 commands from $\AC$ to
\isatt{AC{\isacharunderscore}func{\isacharunderscore}Pow}. This last
one is the
choice principle that we use in our formalization of Pointed $\DC$, and states 
the existence of
choice functions on $\P(A)\sm \{\mty\}$):
\[
\exists (s: \P(A)\sm \{\mty\}\to A). \, \forall X\sbq A.\ X\neq \mty \limp
s(X) \in X.
\]
Another advantage of taking
\isatt{AC{\isacharunderscore}func{\isacharunderscore}Pow} as a
starting point is that it does not involve proper classes: The version
of  $\AC$ in Isabelle/ZF corresponds to an axiom scheme of first order
logic and as such is not a standard formulation. 
%% 26 commands to get Pointed $\DC$.

@@

The proof is mostly routine; after a few lemmas (26 proof
commands in total) we obtain the
following theorem:
